---
title: "Assignment_7"
output: word_document
date: "2025-04-02"
---

Part 1

```{r}
df_powerpose <- read.table("C:/Users/choud/OneDrive - IIT Kanpur/ECO Resources/CGS698C/df_powerpose.csv",header=T,sep=",")
head(df_powerpose)
```

```{r}
library(brms)
library(rstan)
library(dplyr)
```

We need to set the hptreat "Low" as reference and we need to set the priors. I have used high values of sigma so that it explores very well as we have no idea of the priors of the data.

The weakly informative priors we are defining are b1 \~ Normal(0,20) b2 \~Normal (0,20) e \~ uniform (0,20)

```{r}
df_powerpose<-df_powerpose %>% 
  mutate(delta_test=testm2-testm1,
         hptreat=factor(hptreat,levels=c("Low","High")))

priors<-c(prior(normal(0,20),class=Intercept),
          prior(normal(0,20),class=b),
          prior(normal(0,20),class=sigma))

```

```{r}
model <- brm(
  formula = delta_test ~ hptreat, 
  data = df_powerpose,
  family = gaussian(),
  prior=priors,
  chains=4,cores=4,
  iter=2000,warmup=1000)
summary(model)
```

```{r}
library(bayesplot)
mcmc_trace(model)
```

TO check the hypothesis of the hptreatHigh\>0

```{r}
hypothesis(model,"hptreatHigh>0")
```

This proves that that the results align with the research hypothesis. Assigning a subject a high power pose vs. a low power pose will lead to higher testosterone levels after treatment.

Part 2

Exercise 2.1

```{r}
poisson_crossings<-function(sentence_length,alpha,beta){
  log_lambda<-alpha+beta*sentence_length
  lambda_i<-exp(log_lambda)
  crossings<-rpois(1,lambda_i)
  return(crossings)
}
```

Exercise 2.2

```{r}
sentence_length<-4
nsamples<-1000
alpha_samples<-rnorm(nsamples,mean=0.15,sd=0.1)
beta_samples<-rnorm(nsamples,mean=0.25,sd=0.05)
predicted_crossings <- sapply(1:nsamples, function(i) {
  poisson_crossings(sentence_length, alpha_samples[i], beta_samples[i])
})
library(ggplot2)
ggplot(data.frame(predicted_crossings), aes(x = predicted_crossings)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.7, color = "black") +
  labs(title = "Prior Predictive Distribution of Crossing Dependencies",
       x = "Number of Crossings", y = "Frequency") +
  theme_minimal()

```

Exercise 2.3

```{r}
df_crossings<-read.table("C:/Users/choud/OneDrive - IIT Kanpur/ECO Resources/CGS698C/crossings.csv",sep=",",header=T)
df_crossings
```

Convert Language column to categorical variable

```{r}
df_crossings$language_binary <- ifelse(df_crossings$Language == "German", 1, 0)
df_crossings
```

Defining Priors and Fitting the Model M1

```{r}
# Define priors
priors_M1 <- c(
  prior(normal(0.15, 0.1), class = "Intercept"),  # α ~ Normal(0.15, 0.1)
  prior(normal(0, 0.15), class = "b")  # β ~ Normal(0, 0.15)
)

# Fit Poisson model M1
model_M1 <- brm(
  nCross ~ s.length, 
  data = df_crossings, 
  family = poisson(), 
  prior = priors_M1, 
  chains = 4, iter = 2000, warmup = 1000, seed = 123
)

# Summary of M1
summary(model_M1)
```

```{r}
priors_M2 <- c(
  prior(normal(0.15, 0.1), class = "Intercept"),  # α ~ Normal(0.15, 0.1)
  prior(normal(0, 0.15), class = "b"),  # β ~ Normal(0, 0.15)
  prior(normal(0, 0.15), class = "b", coef = "language_binary"),  
  prior(normal(0, 0.15), class = "b", coef = "s.length:language_binary")  
)
```

```{r}
model_M2 <- brm(
  nCross ~ s.length * language_binary, 
  data = df_crossings, 
  family = poisson(), 
  prior = priors_M2, 
  chains = 4, iter = 2000, warmup = 1000, seed = 123
)

# Summary of M2
summary(model_M2)
```

Exercise 2.4

```{r}
observed <- read.table("C:/Users/choud/OneDrive - IIT Kanpur/ECO Resources/CGS698C/crossings.csv",sep=",",header=T)
observed %>% group_by(Language,s.length) %>%
 summarise(mean.crossings=mean(nCross)) %>%
 ggplot(aes(x=s.length,y=mean.crossings,
 group=Language,color=Language))+
 geom_point()+geom_line()
observed$s.length <- observed$s.length- mean(observed$s.length)
observed$lang <- ifelse(observed$Language=="German",1,0)
lpds.m1 <- c()
lpds.m2 <- c()
untested <- observed
for(k in 1:5){
 ytest <- sample_n(untested,size=nrow(observed)/5)
 ytrain <- setdiff(observed,ytest)
 untested <- setdiff(untested,ytest)
 fit.m1<-
   brm(nCross ~ 1 + s.length,data=ytrain,
       family = poisson(link = "log"),
       prior = c(prior(normal(0.15, 0.1), class = Intercept),
                 prior(normal(0, 0.15), class = b)),
       cores=4)
 fit.m2<-
   brm(nCross ~ 1 +s.length+lang+s.length*lang,
       data=ytrain,
       family=poisson(link = "log"),
       prior = c(prior(normal(0.15, 0.1), class = Intercept),
                 prior(normal(0, 0.15), class = b)),
       cores=4)
 post.m1 <- posterior_samples(fit.m1)
 post.m2 <- posterior_samples(fit.m2)
 lppd.m1 <- 0
 lppd.m2 <- 0
 for(i in 1:nrow(ytest)){
  lpd_im1 <- log(mean(dpois(ytest[i,]$nCross,
              lambda=exp(post.m1[,1]+
              post.m1[,2]*ytest[i,]$s.length))))
 lppd.m1 <- lppd.m1 + lpd_im1
 lpd_im2 <- log(mean(dpois(ytest[i,]$nCross,
              lambda=exp(post.m2[,1]+
              post.m2[,2]*ytest[i,]$s.length+
                post.m2[,3]*ytest[i,]$lang+
                post.m2[,4]*ytest[i,]$s.length*ytest[i,]$lang)
              )))
 lppd.m2 <- lppd.m2 + lpd_im2
 }
 lpds.m1 <- c(lpds.m1,lppd.m1)
 lpds.m2 <- c(lpds.m2,lppd.m2)
}
elpd.m1 <- sum(lpds.m1)
elpd.m2 <- sum(lpds.m2)
difference_elpd <- elpd.m2-elpd.m1

```
